How to write a PhD in Biological Sciences: a guide for the uninitiated

### 2021-04-24 00-author.Rmd
Add files via upload
# After some small edits over the week

---
output:
  pdf_document: default
  html_document: default
  word_document: default
---
# About the Author {- #author}

@@

@@

This book is written in bookdown [@xie2016bookdown] specifically to make it a 'live project' that will be open to anyone who wants to contribute, improve, or use as the basis for your own book in a different language, or for a different region on the world. It would be impossible to cover every aspect of writing a PhD in Biological Sciences, and it may be that I have missed out ones that are very important to you. Equally, what is currently written will become obsolete very quickly as new initiatives are begun, and old problems are resolved. For this reason, this guide needs to be a 'living document', and anyone who wants to provide feedback or contribute new sections is more than welcome. Please feel free to open an issue, or make a Pull Request if you spot a typo.

>>>>

---
output:
  html_document: default
  pdf_document: default
  word_document: default
---
# About the Author {- #author}

@@

```{r fig.align='right', echo=FALSE, include=identical(knitr:::pandoc_to(), 'html'), fig.link='http://john.measey.com/About-Me'}
knitr::include_graphics('figures/john.jpg', dpi = NA)
```

@@

This book is written in bookdown [@xie2016bookdown] specifically to make it a 'live project' that will be open to anyone who wants to contribute, improve, or use as the basis for your own book. One of the amazing potentials for bookdown books is that all the files for this book are hosted on [Github](https://github.com/johnmeasey/How-to-write-a-PhD-in-Biological-Sciences/tree/main). You have the opportunity to fork this repository and write your own version for a different discipline, a different language or for a different region of the world. It would be impossible to cover every aspect of writing a PhD in Biological Sciences, and it may be that I have missed out ones that are very important to you. Equally, parts of what is currently written will become obsolete very quickly as new initiatives are begun, and old problems are resolved. For this reason, this guide needs to be a 'living document', and anyone who wants to provide feedback or contribute new sections is more than welcome. Please feel free to open an issue, or make a Pull Request if you spot a typo.

### 2021-04-24 2.08-citations.Rmd
Add files via upload
# After some small edits over the week

^1^Berger L, Speare R, Hyatt A. _Chytrid fungi and amphibian declines: overview, implications and future directions. Declines and disappearances of Australian frogs_ Environment Australia, Canberra. 1999;1999:23–33.

^2^Collins JP, Crump ML, Lovejoy III TE. _Extinction in our times: global amphibian decline_. Oxford University Press; 2009.

^3^Daszak P, Cunningham AA, Hyatt AD. Infectious disease and amphibian population declines. _Diversity and Distributions_. 2003;9(2):141–50.

^4^Dufresnes C, Dubey S, Ghali K, Canestrelli D, Perrin N. Introgressive hybridization of threatened European tree frogs (*Hyla arborea*) by introduced _H. intermedia_ in Western Switzerland. _Conservation Genetics_. 2015;16(6):1507–13.

^5^Kupferberg SJ. Bullfrog (*Rana catesbeiana*) invasion of a California river: the role of larval competition. _Ecology_. 1997;78(6):1736–51.

^6^La Marca E, Lips KR, Lotters S, Puschendorf R, Ibanez R, Rueda-Almonacid JV, et al. Catastrophic population declines and extinctions in Neotropical harlequin frogs (Bufonidae: *Atelopus*) _Biotropica: The Journal of Biology and Conservation_. 2005;37(2):190–201.

^7^Martel A, Spitzen-van der Sluijs A, Blooi M, Bert W, Ducatelle R, Fisher MC, et al. *Batrachochytrium salamandrivorans* sp. nov. causes lethal chytridiomycosis in amphibians. _Proceedings of the National Academy of Sciences_. 2013;110(38):15325–9.

^8^Pimm SL, Jenkins CN, Abell R, Brooks TM, Gittleman JL, Joppa LN, et al. The biodiversity of species and their rates of extinction, distribution, and protection. _Science_. 2014;344(6187):1246752.

^9^Wake DB, Vredenburg VT. Are we in the midst of the sixth mass extinction? A view from the world of amphibians. _Proceedings of the National Academy of Sciences_. 2008;105(Supplement 1):11466–73.

@@

Berger, L., Speare, R., Hyatt, A., 1999. _Chytrid fungi and amphibian declines: overview, implications and future directions. Declines and disappearances of Australian frogs_. Environment Australia, Canberra 1999, 23–33.

Collins, J.P., Crump, M.L., Lovejoy III, T.E., 2009. _Extinction in our times: global amphibian decline_. Oxford University Press.

Daszak, P., Cunningham, A.A., Hyatt, A.D., 2003. Infectious disease and amphibian population declines. _Diversity and Distributions_ 9, 141–150.

Dufresnes, C., Dubey, S., Ghali, K., Canestrelli, D., Perrin, N., 2015. Introgressive hybridization of threatened European tree frogs (*Hyla arborea*) by introduced _H. intermedia_ in Western Switzerland. _Conservation Genetics_ 16, 1507–1513.

Kupferberg, S.J., 1997. Bullfrog (*Rana catesbeiana*) invasion of a California river: the role of larval competition. _Ecology_ 78, 1736–1751.

La Marca, E., Lips, K.R., Lotters, S., Puschendorf, R., Ibanez, R., Rueda-Almonacid, J.V., Schulte, R., Marty, C., Castro, F., Manzanilla-Puppo, J., others, 2005. Catastrophic population declines and extinctions in Neotropical harlequin frogs (Bufonidae: *Atelopus*) _Biotropica: The Journal of Biology and Conservation_ 37, 190–201.

Martel, A., Spitzen-van der Sluijs, A., Blooi, M., Bert, W., Ducatelle, R., Fisher, M.C., Woeltjes, A., Bosman, W., Chiers, K., Bossuyt, F., others, 2013. *Batrachochytrium salamandrivorans* sp. nov. causes lethal chytridiomycosis in amphibians. _Proceedings of the National Academy of Sciences_ 110, 15325–15329.

Pimm, S.L., Jenkins, C.N., Abell, R., Brooks, T.M., Gittleman, J.L., Joppa, L.N., Raven, P.H., Roberts, C.M., Sexton, J.O., 2014. The biodiversity of species and their rates of extinction, distribution, and protection. _Science_ 344, 1246752

Wake, D.B., Vredenburg, V.T., 2008. Are we in the midst of the sixth mass extinction? A view from the world of amphibians. _Proceedings of the National Academy of Sciences_ 105, 11466–11473.

@@

Another common mistake is to forget which paper has which information. You can try to make sure that you don’t do this by taking better notes or a more accurate plan.

>>>>

The corresponding references in [Vancouver](vancouver.html) format are in the section on [references](references.html).

@@

You can find the corresponding references in [Harvard](harvard.html) format in the [references](references.html) section.

@@

Another common mistake is to forget which paper has which information. You can try to make sure that you don’t do this by taking better notes or a more accurate plan. And do the citations before or as you write, not afterwards!

### 2021-04-24 2.09-databases.Rmd
Add files via upload
# After some small edits over the week

### Citations

@@

#### What else are literature databases used for?

>>>>

## Moving items into your reference manager

An invaluable way in which databases have become extremely useful is the ability to move items that you have found during your search directly into your [reference manager](referencesoftware.html). Beware that these regularly come with errors, and that the amount of errors is related to your database source (Scopus and WoS are much better than Google Scholar). This means that after importing the refrences, you will need to go back to your reference manager and edit them or they will come out incorrectly when you use them. Of particular note to biologists is the need to have all species names in [italics](taxonomy.html), and you'll need to do this yourself.

## Citations

@@

### What else are literature databases used for?

### 2021-04-24 2.10a-sortware.Rmd
Add files via upload
# After some small edits over the week

If you feel comfortable with a word processing package with a nice GUI, like Microsoft Word, then go ahead and use it. My best suggestion is that you do not save your files in their default .docx format. Instead make sure that all of your files are saved as .rtf or html. This should make little difference to your opening and editing but will mean that all of your texts are forward compatible with other kinds of software. It will reduce your file size and reduce the chances of file corruption: i.e. when part of the file malfunctions and you lose the contents. That’s right, in addition to loss of data by back up deficiencies [see Part III](data.html), you also have the risk of having corrupt files that will also mean you lose all of the contents. Even if you always use Word and never use any other file format, do yourself a favour and save each version as a .rtf or .txt file as well. Just like a back up, this means that you won’t have lost all of the text if a file goes corrupt. And yes, I do know people who have multiple backup versions of corrupt files: it’s happened to me.

>>>>

Another important consideration is whether or not your chosen word processing package will allow seemless integration with your [reference manager](referencesoftware.html). Your lab may have a communal repository of references, and so this may then constrain what you are comfortable using.

If you feel comfortable with a word processing package with a nice GUI, like Microsoft Word, then go ahead and use it. These certainly have advantages like being able to get a synonym at the click of a mouse (Figure \@ref(fig:synonym)). My best suggestion is that you do not save your files in their default ".docx" format. Instead make sure that all of your files are saved as .rtf or html. This should make little difference to your opening and editing but will mean that all of your texts are forward compatible with other kinds of software. It will reduce your file size and reduce the chances of file corruption: i.e. when part of the file malfunctions and you lose the contents. That’s right, in addition to loss of data by back up deficiencies [see Part III](data.html), you also have the risk of having corrupt files that will also mean you lose all of the contents. Even if you always use Word and never use any other file format, do yourself a favour and save each version as a .rtf or .txt file as well. Just like a back up, this means that you won’t have lost all of the text if a file goes corrupt. And yes, I do know people who have multiple backup versions of corrupt files: it’s happened to me.

Keeping files in a generic (rather than software manufacturer's) format will allow easy sharing of these files, and access to them in future. It might seem inconceivable now that anyone would not be able to read a .docx file, but in 10 years, you may be so used to the latest format that after another 5 years it becomes very difficult to retrieve information from a .docx file without paying someone to retrieve your data.

### 2021-04-24 2.22-transparency.Rmd
Add files via upload
# After some small edits over the week

## Starting out transparent {#transparency}

@@

In this book, I have tried to consistently encourage you toward transparent and open practices in science. Here my aim is to introduce this topic and provide an overview of the main areas in which you can currently make a difference by opening up your research. As the Centre for Open Science explains (COS: [see here](https://youtu.be/9YuNGB3vNOw)), this will take a shift in the culture of the scientific community, which is why you need to understand and adopt a transparent approach from the outset of your research career. There are much better and more comprehensive guides and online resources so I encourage you to look through the literature cited here. Although I advocate hard for the COS, please be aware that this is not the only set of guidelines (see also [recent moves by the Public Library of Science](https://plos.org/open-science/preregistration/)).

### Preregistration

@@

#### Does this mean that you can’t perform any data exploration?

@@

#### Will preregistration of research eliminate the bias from science?

It is probably too early to tell [@rubin2020does], but it is certainly a good place to start [@nosek2018preregistration]. The more researchers know and subscribe to [transparency](transparency2.html) in their research, the more it will shift the culture in science for the better [@forstmeier2017detecting].However, there are more biases existing in science, and the system needs buy-in from stakeholders to promote equity by eliminaion of bias.

#### What Platform should you use?

@@

### Transparency as you move forwards

>>>>

# Starting out transparent {#transparency}

@@

In this book, I have tried to consistently encourage you toward transparent and open practices in science. Here my aim is to introduce this topic and provide an overview of the main areas in which you can currently make a difference by opening up your research. As the Centre for Open Science explains (COS: [see here](https://youtu.be/9YuNGB3vNOw)), this will take a shift in the culture of the scientific community, which is why you need to understand and adopt a transparent approach from the outset of your research career. There are much better and more comprehensive guides and online resources so I encourage you to look through the literature cited here. Although I advocate hard for the COS, please be aware that this is not the only set of guidelines (see also [recent moves by the Public Library of Science](https://plos.org/open-science/preregistration/)) and [Peer Community In](https://peercommunityin.org/).

## Preregistration

@@

### Does this mean that you can’t perform any data exploration?

@@

### Will preregistration of research eliminate the bias from science?

It is probably too early to tell [@rubin2020does], but it is certainly a good place to start [@nosek2018preregistration]. The more researchers know and subscribe to [transparency](transparency2.html) in their research, the more it will shift the culture in science for the better [@forstmeier2017detecting].However, there are more biases existing in science, and the system needs buy-in from stakeholders to promote equity by elimination of bias.

A new initiative from [Peer Community In](https://peercommunityin.org/) is the possibility of submitting to their [Registered Reports](https://rr.peercommunityin.org/), which goes much further towards removing the confirmation bias. The Registered Report (RR) is in effect the registration of a proposal (i.e. preregistration) with review. If the RR is approved by reviewers then the study is, in principle, given the green light for publication whether or not the hypothesis posed is accepted or rejected. I say "in principle" because those same reviewers are shown the ms again once the results are in. They need to check that the methods proposed were followed, the analyses were conducted in the same way they were proposed, and that the conclusions are justified by the results. In terms of figure \@ref(fig:preregistration), Peer Community In are offering to organise the two yellow review arrows. In addition, there are a bunch of journals that have already signed up to accept RRs that are signed off after completion (notable among these is [PeerJ](www.peerj.com)). To me, this represents an important step in the right direction toward transparency, and the elimination of confirmation bias. What would be great to see is the number of conventional journals sign up with RRs based on the quality of the study design and execution, and the concomitant abandonment of Impact Factor as a driving force in publishing.

### What Platform should you use?

@@

## Transparency as you move forwards

### 2021-04-24 3.08-References.Rmd
Add files via upload
# After some small edits over the week

Putting together the reference section of your chapter or paper can take a very long time. There are quite a lot of different software packages out there to help you do this (e.g. [Endnote](https://endnote.com/), [Mendelay](https://www.mendeley.com/guides/desktop), [Zotero](https://www.zotero.org/) - see more about these [here](https://subjectguides.library.american.edu/c.php?g=479020&p=3323781)). These are a mixed blessing. Firstly, most of them are not free, so unless your university subscribes to their providers you would have to pay for the privilege of using them. Next, all of the ones that I've ever used are very demanding and do not necessarily work well with your word processor of choice. Given all the time I have invested in reference formatting software, I think I might have been better off just copying and pasting references I have at the end of a document and then formatting them by hand. This tends to be how I do it now. Having said this, I used Zotero as the referencing software for this book. It took very little time to learn how to use, it was painless to enter (most) of the references into the database, and it is free and works with [Bookdown](bookdown.org) and Google Docs.

Earlier in this book, I provided a guide to the differences between Vancouver and

@@

Lots of journals are very fussy about the way that citations are given in their publications. I think it's really very cheeky especially as many of them also force you to pay for the privilege of doing all the work. There are some exceptions to this (e.g. [PeerJ](www.peerj.com)).

There are lots of better things that you could be doing with your time instead of formatting references. And so I'm not going to spend an awful long time here telling you that you must do it right. I really believe that on first submission of your manuscript you should be able to submit your references in any format. it really only makes sense for you to do all the work formatting them if you aren't going to pay to have published the paper.

There are lots of anal academics out there who will delight in looking for every missing full-stop, comma and capital-letter missing from your references. Perhaps it's because they feel secure pointing out these errors, but are incapable of knowing whether or not you've written good or bad science. Please don't join them and if you can look for a better world where you spend your valuable time doing more important things. For me, at this moment, it does not include any referencing software that I know of.

>>>>

@@

[Earlier in this book](citations.html), I provided a guide to the differences between Vancouver and Harvard referencing styles. Here I provide the output for the references that were given in that section to demonstrate the way in which these different referencing styles take up very different amounts of space. In that section, I provided some quotes from a published paper that had 9 citations. Below, you can see what these 9 citations look like in the two different generic outputs: Vancouver and Harvard.

## Vancouver style {#vancouver}
The following references follow the text written in [Vancouver style](citations.html) in the section on citations. Note that the order of citations here corresponds to the order in which they are cited.

>^1^Wake DB, Vredenburg VT. Are we in the midst of the sixth mass extinction? A view from the world of amphibians. _Proceedings of the National Academy of Sciences_. 2008;105(Supplement 1):11466–73.
>
>^2^Collins JP, Crump ML, Lovejoy III TE. _Extinction in our times: global amphibian decline_. Oxford University Press; 2009.
>
>^3^Pimm SL, Jenkins CN, Abell R, Brooks TM, Gittleman JL, Joppa LN, et al. The biodiversity of species and their rates of extinction, distribution, and protection. _Science_. 2014;344(6187):1246752.
>
>^4^Kupferberg SJ. Bullfrog (*Rana catesbeiana*) invasion of a California river: the role of larval competition. _Ecology_. 1997;78(6):1736–51.
>
>^5^Dufresnes C, Dubey S, Ghali K, Canestrelli D, Perrin N. Introgressive hybridization of threatened European tree frogs (*Hyla arborea*) by introduced _H. intermedia_ in Western Switzerland. _Conservation Genetics_. 2015;16(6):1507–13.
>
>^6^Berger L, Speare R, Hyatt A. _Chytrid fungi and amphibian declines: overview, implications and future directions. Declines and disappearances of Australian frogs_ Environment Australia, Canberra. 1999;1999:23–33.
>
>^7^Daszak P, Cunningham AA, Hyatt AD. Infectious disease and amphibian population declines. _Diversity and Distributions_. 2003;9(2):141–50.
>
>^8^La Marca E, Lips KR, Lotters S, Puschendorf R, Ibanez R, Rueda-Almonacid JV, et al. Catastrophic population declines and extinctions in Neotropical harlequin frogs (Bufonidae: *Atelopus*) _Biotropica: The Journal of Biology and Conservation_. 2005;37(2):190–201.
>
>^9^Martel A, Spitzen-van der Sluijs A, Blooi M, Bert W, Ducatelle R, Fisher MC, et al. *Batrachochytrium salamandrivorans* sp. nov. causes lethal chytridiomycosis in amphibians. _Proceedings of the National Academy of Sciences_. 2013;110(38):15325–9.
>
>

## Harvard style {#harvard}
The references below are exactly the same as those above, but follow the [Harvard](citations.html) referencing format. The order of these references is alphabetical. If a paper has the same authors the oldest one usually comes before subsequent papers.

>Berger, L., Speare, R., Hyatt, A., 1999. _Chytrid fungi and amphibian declines: overview, implications and future directions. Declines and disappearances of Australian frogs_. Environment Australia, Canberra 1999, 23–33.
>
>Collins, J.P., Crump, M.L., Lovejoy III, T.E., 2009. _Extinction in our times: global amphibian decline_. Oxford University Press.
>
>Daszak, P., Cunningham, A.A., Hyatt, A.D., 2003. Infectious disease and amphibian population declines. _Diversity and Distributions_ 9, 141–150.
>
>Dufresnes, C., Dubey, S., Ghali, K., Canestrelli, D., Perrin, N., 2015. Introgressive hybridization of threatened European tree frogs (*Hyla arborea*) by introduced _H. intermedia_ in Western Switzerland. _Conservation Genetics_ 16, 1507–1513.
>
>Kupferberg, S.J., 1997. Bullfrog (*Rana catesbeiana*) invasion of a California river: the role of larval competition. _Ecology_ 78, 1736–1751.
>
>La Marca, E., Lips, K.R., Lotters, S., Puschendorf, R., Ibanez, R., Rueda-Almonacid, J.V., Schulte, R., Marty, C., Castro, F., Manzanilla-Puppo, J., others, 2005. Catastrophic population declines and extinctions in Neotropical harlequin frogs (Bufonidae: *Atelopus*) _Biotropica: The Journal of Biology and Conservation_ 37, 190–201.
>
>Martel, A., Spitzen-van der Sluijs, A., Blooi, M., Bert, W., Ducatelle, R., Fisher, M.C., Woeltjes, A., Bosman, W., Chiers, K., Bossuyt, F., others, 2013. *Batrachochytrium salamandrivorans* sp. nov. causes lethal chytridiomycosis in amphibians. _Proceedings of the National Academy of Sciences_ 110, 15325–15329.
>
>Pimm, S.L., Jenkins, C.N., Abell, R., Brooks, T.M., Gittleman, J.L., Joppa, L.N., Raven, P.H., Roberts, C.M., Sexton, J.O., 2014. The biodiversity of species and their rates of extinction, distribution, and protection. _Science_ 344, 1246752
>
>Wake, D.B., Vredenburg, V.T., 2008. Are we in the midst of the sixth mass extinction? A view from the world of amphibians. _Proceedings of the National Academy of Sciences_ 105, 11466–11473.
>
>

Be aware that these styles (Vancouver and Harvard) represent two broad approaches, every publisher has their own variation on these, and that they can vary substantially between different journals. Full details are always provided by the journal (in their **Instructions to Authors**), and you can refer to these to see how your references need to be formatted. I also provide some pointers towards using [reference managers](referencesoftware.html). These can save you a lot of time, if you've already invested time in setting them up. Alternatively, they can be very frustrating.

Lots of journals (or perhap's it's their editors) are very fussy about the way that citations are given in their publications. If you have used a decent reference manager then you won't care, and it'll just be a case of finding the appropriate format from their repository (or making it yourself).

There are lots of better things that you could be doing with your time instead of formatting references. And so I'm not going to spend a long time here telling you that you must do it right. I really believe that on first submission of your manuscript you should be able to submit your references in any format. it really only makes sense for you to do all the work formatting them if you aren't going to pay to have published the paper. An increasing number of journals do allow you to use any referencing style (within the generic Vancouver or Harvard format) when you first submit. You are only then required to format them if your ms is likely to be accepted.

There are lots of anal academics out there who will delight in looking for every missing full-stop, comma and capital-letter missing from your references. Perhaps it's because they feel secure pointing out these errors, but are incapable of knowing whether or not you've written good or bad science. Please don't join them and if you can look for a better world where you spend your valuable time doing more important things.

### 2021-04-24 index.Rmd
Add files via upload
# After some small edits over the week

--- 
title: 'How to write a PhD in Biological Sciences: a guide for the uninitiated'
author: "John Measey"
date: "`r Sys.Date()`"
documentclass: krantz
bibliography:
- book.bib
- packages.bib
biblio-style: apalike
link-citations: yes
colorlinks: yes
lot: yes
lof: yes
site: bookdown::bookdown_site
description: You don’t have to be a genius to write a PhD. It will always involve
  a lot of hard work and dedication, but the process of writing up is a whole lot
  easier if you understand the basic ground rules. This book is a guide specifically
  for writing a PhD in the *biological sciences*. It will guide you from deciding
  whether or not you should actually do a PhD, provide practical, up-to-the-minute
  guides to getting started with planning and writing, and consolidate this with the
  nuts and bolts of writing for the biological sciences. The book guides you through
  the do’s and don’ts of writing a thesis, and taking your thesis chapters to publication
  in peer-reviewed journals. Whether you are a student just starting your studies,
  or a supervisor struggling to cope, this book will provide the insider information
  you need to get ahead.
github-repo: yihui/bookdown-crc
graphics: yes
---

@@

@@

This book started as a series of blog posts written by me for my students (john.measey.com/blog). I wrote them because I felt that I was repeating the same advice year after year. Instead of rewriting comments on manuscripts time after time, I thought I would consolidate them and put them all in one place. Later, I started looking for particular problems that students were having, and then writing blog posts to try to preempt those problems, and to provide solutions before tasks were completed. Finally, I have tried to fill in the missing gaps in the blog posts. But there are still plenty of gaps, and I regard this book as a work-in-progress. It is my intention that this book will always be added to, and improved by a community of practice, and I hope that this will include you.

@@

Before I go on, there are a great many people that I need to thank. First and foremost are my students, past and present, who have inspired me to put together first the blog posts and then the book. It is because you wanted more that I put this together. I have also been a student, and have been inspired by colleagues around the world who have been exemplary advisors. This book contains lots of links to blogs and articles written and posted freely on the internet by others who also aim to demystify and help. I thank this greater academic community for sharing and inspiring. Thanks go to the many reviewers and editors who have taken their time to improve my writing. I am still learning. Lots of the text in this book has been improved by feedback from my students and postdocs. A special mention must go to my brother, Richard, who has hosted my lab website for more than a decade, and especially for saving blog posts from hacking attacks. Thanks also to my wife, Thalassa, who proofread many of the blog posts after I had published them late at night, so that I could correct them over breakfast in the morning. James Baxter-Gilbert, Anthony Herrel,  Lisa O'Connell, Mark O'Connell, James Vonesh all read parts and commented. Thanks are also due to my colleagues at the Centre for Invasion Biology, the Department of Botany and Zoology, and Stellenbosch University. A special thanks to the librarians who have supported many of my more extreme rantings about publishers.

>>>>

--- 
title: 'How to write a PhD in Biological Sciences: a guide for the uninitiated'
author: "John Measey"
date: "`r Sys.Date()`"
documentclass: krantz
bibliography:
- book.bib
- packages.bib
biblio-style: apalike
link-citations: yes
colorlinks: yes
lot: yes
lof: yes
site: bookdown::bookdown_site
description: You don’t have to be a genius to write a PhD. It will always involve
  a lot of hard work and dedication, but the process of writing up is a whole lot
  easier if you understand the basic ground rules. This book is a guide specifically
  for writing a PhD in the biological sciences. It will guide you from deciding
  whether or not you should actually do a PhD, provide practical, up-to-the-minute
  guides to getting started with planning and writing, and consolidate this with the
  nuts and bolts of writing for the biological sciences. The book guides you through
  the do’s and don’ts of writing a thesis, and taking your thesis chapters to publication
  in peer-reviewed journals. Whether you are a student just starting your studies,
  or a supervisor struggling to cope, this book will provide the insider information
  you need to get ahead.
github-repo: yihui/bookdown-crc
graphics: yes
---

@@

# Under Contract {-}
This book is currently under contract with CRC Press. 
```{r fig.align='center', echo=FALSE, include=identical(knitr:::pandoc_to(), 'html'), fig.link='http://john.measey.com/Writing_Publishing'}
knitr::include_graphics('figures/Cover.png', dpi = NA)
```

@@

This book started as a series of blog posts written by me for my students ([john.measey.com/blog](john.measey.com/blog)). I wrote them because I felt that I was repeating the same advice year after year. Instead of rewriting comments on manuscripts time after time, I thought I would consolidate them and put them all in one place. Later, I started looking for particular problems that students were having, and then writing blog posts to try to preempt those problems, and to provide solutions before tasks were completed. Finally, I have tried to fill in the missing gaps in the blog posts. But there are still plenty of gaps, and I regard this book as a work-in-progress. It is my intention that this book will always be added to, and improved by a community of practice, and I hope that this will include you.

@@

Before I go on, there are a great many people that I need to thank. First and foremost are my students, past and present, who have inspired me to put together first the blog posts and then the book. It is because you wanted more that I put this together. I have also been a student, and have been inspired by colleagues around the world who have been exemplary advisors. This book contains lots of links to blogs and articles written and posted freely on the internet by others who also aim to demystify and help. I thank this greater academic community for sharing and inspiring. Thanks go to the many reviewers and editors who have taken their time to improve my writing. I am still learning. Lots of the text in this book has been improved by feedback from my students and postdocs. A special mention must go to my brother, Richard, who has hosted my lab website for more than a decade, and especially for saving blog posts from hacking attacks. Thanks also to my wife, Thalassa, who proofread many of the blog posts after I had published them late at night, so that I could correct them over breakfast in the morning. James Baxter-Gilbert, Anthony Herrel,  Allan Ellis, James Vonesh all read parts and commented. Thanks are also due to my colleagues at the Centre for Invasion Biology, the Department of Botany and Zoology, and Stellenbosch University. A special thanks to the librarians who have supported many of my more extreme rantings about publishers.
